Hydras \& Co.: Formalized mathematics in Coq\\ for inspiration and entertainment

### 2022-06-25 doc/JFLA2022/Slides/Slides.tex
last changes in Slides ?

  {

 \textbf{Definition:} \textcolor{mathcolor}{$H'_\alpha(i)$} is the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (cf. previous slide).

 {\color{mathcolor}
\begin{align}
  H'_0(i) & = i\\
  H'_{\beta+1}(i) &=H'_\beta(i+1) \\
  H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{($\alpha$ limit ordinal)} 
\end{align}}}

{$H'$ is a slight variant of the \emph{Hardy Hierarchy}. In \coq, we implement it with the \texttt{equations} plug-in.}
%\vspace{6pt}

@@

% {\color{mathcolor}
% \begin{align}
% F_0(i)& =i+1\\
% F_{\beta+1}(i)&= (F_\beta)^{(i+1)}(i)\\
% F_\alpha(i) &= F_{(\canonseq{\alpha}{i})} (i) \quad \textit{($\alpha$ limit ordinal)}
% \end{align}}}
\end{block}
\end{frame}

@@

       \item     If \textcolor{mathcolor}{$\alpha \rplus{i,i,\dots,i} \beta$} then \textcolor{mathcolor}{$H'_\beta(i)\leq H'_\alpha(i)$}.

@@

 \begin{block}{Examples}

@@

        \begin{align*}
        H'_{\omega^\omega}(k)=\,& \textbf{let}\,F\;f\;i := f^{(i+1)}(i)\,
                                  \textbf{in}\, F^{(k)}(\lambda k.\,2k+1)\,k\\
      H'_{\omega^3}(k) \geq\, & \textit{hyper\_exp2}(k+1) \\
 %     H'_{\omega^\alpha}(k)\geq\,& F_\alpha(k)\; (k>0)\\
    \end{align*}
       }%mathcolor
  \end{block}

@@

\begin{frame}[label={fgftheorems}]
\label{thms-fastgrowing}
  \begin{block}{Theorem}
    % \begin{itemize}
    % \item \textcolor{mathcolor}{$F_\alpha$} is primitive recursive iff \textcolor{mathcolor}{$\alpha<\omega$}.
    %  \item  
    The function which computes the length of hydra battles
    for \textcolor{mathcolor}{$\alpha$} (according to the initial index) is not primitive recursive
    (for $\textcolor{mathcolor}{\alpha\geq \omega^\omega}$).
  %  \end{itemize}

  \begin{block}{Sketch of proof:}

   We apply monotony properties and the famous lemma:
    \vspace{5pt}

       \begin{quote}
      {\color{mathcolor}
        For any $n$-ary primitive recursive  function $f$, 
        there exists some $q$ such that the following inequality holds:
         \[\forall\,x_1,\dots,x_n,  
          f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))\]}
    \end{quote}
  %  (Proof by induction on any primitive recursive function of any arity.)
  \end{block}

  \end{block}
\end{frame}

>>>>

\begin{block}{Definition:}
  \textcolor{mathcolor}{$H'_\alpha(i)$} is the least $l$ such that \textcolor{mathcolor}{$\alpha_l=0$} (cf. previous slide).

 {\color{mathcolor}
\begin{align}
  H'_0(i) & = i\\
  H'_{\beta+1}(i) &=H'_\beta(i+1) \\
  H'_\alpha(i) &= H'_{(\canonseq{\alpha}{i+1})}(i)  \quad\textit{($\alpha$ limit ordinal)} 
\end{align}}

$H'$ is a slight variant of the \emph{Hardy Hierarchy}. In \coq, we implement it with the \texttt{equations} plug-in.
\end{block}

\begin{block}{Example:}
  {\color{mathcolor}
  $$      H'_{\omega^\omega}(i)=\, \textbf{let}\,F\;f\;k := f^{(k+1)}(k)\,
                                  \textbf{in}\, F^{(i)}(\lambda k.\,2k+1)\,i$$
}
\end{block}

%\vspace{6pt}

@@

% {\color{mathcolor}
% \begin{align}
% F_0(i)& =i+1\\
% F_{\beta+1}(i)&= (F_\beta)^{(i+1)}(i)\\
% F_\alpha(i) &= F_{(\canonseq{\alpha}{i})} (i) \quad \textit{($\alpha$ limit ordinal)}
% \end{align}}}

@@

       \item     If \textcolor{mathcolor}{$\alpha \rplus{i,i,\dots,i} \beta$} then \textcolor{mathcolor}{$H'_\alpha(i)\geq H'_\beta(i)$}.

@@

 \begin{block}{Examples:}

@@

        \begin{align*}
      H'_{\omega^3}(i) \geq\, & \textit{hyper\_exp2}(i+1) \\
      H'_{\omega^\omega}(i)\geq\,& \textit{Ack}(i,i)\; (i\geq 2)\\
    \end{align*}
  }%mathcolor
 Thus, for any \textcolor{mathcolor}{$\alpha\geq\omega^\omega$}, \textcolor{mathcolor}{$H'_\alpha$} is not primitive recursive.
  \end{block}

@@

% \begin{frame}[label={fgftheorems}]
% \label{thms-fastgrowing}
%   \begin{block}{Theorem}
%     % \begin{itemize}
%     % \item \textcolor{mathcolor}{$F_\alpha$} is primitive recursive iff \textcolor{mathcolor}{$\alpha<\omega$}.
%     %  \item  
%     The function which computes the length of hydra battles
%     for \textcolor{mathcolor}{$\alpha$} (according to the initial index) is not primitive recursive
%     (for $\textcolor{mathcolor}{\alpha\geq \omega^\omega}$).
%   %  \end{itemize}

%   \begin{block}{Sketch of proof:}

%    We apply monotony properties and the famous lemma:
%     \vspace{5pt}

%        \begin{quote}
%       {\color{mathcolor}
%         For any $n$-ary primitive recursive  function $f$, 
%         there exists some $q$ such that the following inequality holds:
%          \[\forall\,x_1,\dots,x_n,  
%           f(x_1,\dots,\,x_n)\leq\textrm{Ack}(q,\textrm{max}(x_1,\dots,x_n))\]}
%     \end{quote}
%   %  (Proof by induction on any primitive recursive function of any arity.)
%   \end{block}

%   \end{block}
% \end{frame}
